---
title: "AFA - 4th Session"
author: "Mathis Mourey"
date: "28 octobre 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pricing Option 


```{r packages, echo=T, include=F}
library(ggplot2)
library(quantmod)
library(reshape)
library(arrayhelpers)

```


We starts today's lecture by pricing exotic options using Monte-Carlo procedure. MC for pricing options consists in the following steps: 

1. Choose a process for the underlying asset. Usually, we go for a Geometric Brownian Motion (GBM).
2. Simulate *n* price paths. In other words simulate the full development of the price from $t=0$ to $t=T$, *n* number of times.
3. Compute the payoff of the option at maturity (or $t=T$) for each simulation. 
4. Average the payoff and discount it, in order to bring back the value to present days.

We start by choosing the process for our underlying asset. We choose a common GBM, such that the returns of our underlying follows: 

$$ \frac{dS_t}{S_t} = \mu * dt + \sigma * dB_t  $$
We solve the PDE and we end up with the following closed-form: 

$$ S_{t + dt} = S_t  e^{(\mu - \frac{\sigma^2}{2})dt + \sigma \epsilon \sqrt{dt}} $$
Where, $\epsilon$ is Standard White noise (eq. follows a Standard Gaussian (0, 1)). 

In order to compute $\mu$ (the mean return) and $\sigma$ (the standard deviation of returns), we take the stock price of FB. Then we simulate `nb_mc` number of simulations.

```{r Brownian_motion}

nb_mc = 50
df_mc = data.frame(matrix(NA, nrow=500, ncol=nb_mc))
x = getSymbols('FB', env=NULL )


#take only the 500 last points of the data
df_mc[,1]= tail(index(x), 500)
df_mc[,2] = tail(x[, 6], 500)


#get the mean and std of returns
mu= mean(diff(log(df_mc[,2])), na.rm=T)
sigma = sd(diff(log(df_mc[,2])), na.rm=T)

for(j in 3:(ncol(df_mc)-1)){
  df_mc[249,j] = df_mc[249,2]
  for(i in 250:500){
    df_mc[i, j] = df_mc[i-1,j]*exp((mu-(sigma**2)/2)+sigma*rnorm(1,0,1))
  }
} 

df_mc[,nb_mc] = apply(X=df_mc[,3:(nb_mc-1)], FUN='mean', MARGIN = 1)
tail(df_mc)

```

By using the function `melt()` we allow to create the following graph that integrates all simulated future prices along with the actual one. 

```{r plot}

new_df = melt(df_mc, id='X1')

p = ggplot() + geom_line(data=new_df, aes(x=X1, y=value, colour=variable))
p

```

We develop our program to price exotic option and price an example of Lookback option: 

$$  Payoff = max(S_T - \frac{1}{T} \sum \limits_{t=1}^{T} S_t; 0 )$$


```{r Lookback_pricer}


Lookback_opt_pricer <- function(S, rf, t, sigma, option_type, nb_simu, time_step){
  dt = t/time_step
  drift = (rf-(sigma**2)/2)*dt
  shock = sigma*sqrt(dt)
  
  LB_Brownian = data.frame(matrix(nrow=time_step, ncol=nb_simu))
  LB_Brownian[1,] = rep(S, nb_simu)
  
  for(j in 1:nb_simu ){
    for(i in 2:time_step){
      LB_Brownian[i,j] = LB_Brownian[i-1,j]*exp(drift+shock*rnorm(1))

    }
  }
  
  pre_LB = colMeans(LB_Brownian)
  ST = LB_Brownian[time_step,]
  
  if(option_type=='call'){
    temp = pmax((ST-pre_LB),0)
    LB = sum(temp)/length(temp)
  } else {
    temp= pmax((pre_LB - ST),0)
    LB = sum(temp)/length(temp)
  }
  return(LB*(1+rf)**(-t))
}


Lookback_opt_pricer(S=100, rf=0.04, t=1, sigma=0.1, option_type='put', nb_simu=100, time_step=50)
```

